
# This code event marks baseball pitching throwing trials
# It extracts variable values at all events and saves them as a .csv file
# It overwrites the original .txt file with the updated events file

```{r Load the Data & Event Mark}

rm(list = ls())

# Let the user select the file to import
file_path <- file.choose()

# Read the data
data <- read.delim(file_path, header = TRUE, skip = 8, check.names = FALSE)

# Remove column "49" as it's empty
library(dplyr)
data <- data %>%
  select(! 14) # Remove the final frame in the data if it's empty

# Define sampling rate (Hz)
SampleRate <- 238.095 # change as necessary

# Create and events column and populate with 0s
data$Events <- 0

# Create the column "Time (s)" (divide by sampling rate)
data$`Time (s)` <- data$`Sample #` / SampleRate

library(dplyr)

# Move the 'Events column to the first position and 'Time (s)' to the second position
data <- data %>% 
  select(Events, `Time (s)`, everything())

rm(SampleRate)


# Need to add stride length (% height) here

# Front Foot Contact (FFC)
FFC_condition <- data$`Event - FFC` > 30  # define FFC condition

# Find the index of the first occurrence
FFC_index <- which(FFC_condition)[1]

# Mark the first frame where the condition is met with '1'
if (!is.na(FFC_index)) {
  data$Events[FFC_index] <- 1
}

rm(FFC_condition)


# Find the maximum value of Event - INI
max_INI <- max(data$`Event - INI`, na.rm = TRUE)

# Calculate the starting index for subsetting
start_index <- max(1, which(data$`Event - INI` == max_INI) - 150) # 75 when at 100Hz

# Subset the data from start_index to the end
subset_data <- data[start_index:nrow(data), ]

# Find the index of the first occurrence where the difference between values exceeds 0.0001 within the subset
INI_index <- which(diff(subset_data$`Event - INI`) > 0.0001)[1]

# If INI_index is not NA, adjust the index to reflect the original data
if (!is.na(INI_index)) {
  INI_index <- start_index + INI_index - 1
}

# Mark the first frame where the condition is met with '1' in the original data
if (!is.na(INI_index)) {
  data$Events[INI_index] <- 1
}

rm(max_INI, start_index)

# Maximum External Rotation (MER) <- will be MAW going forward
MER_condition <- data$`Event - MER` == min(data$`Event - MER`, na.rm = TRUE)  # define MER condition

# Find the index of the minimum value occurrence
MER_index <- which(MER_condition)[1]

# Mark the frame where the minimum value occurs with '1'
if (!is.na(MER_index)) {
  data$Events[MER_index] <- 1
}

rm(MER_condition)


# Ball Release (BR)

# Subset the data from FFC_index to the end of the dataset
subset_data <- data[FFC_index:nrow(data), ]

# Find the maximum value of Event - BR within the subset data
max_BR_value <- max(subset_data$`Event - BR`, na.rm = TRUE)
BR_condition <- subset_data$`Event - BR` == max_BR_value # define BR condition

# Find the index of the maximum value of Event - BR within the subset data
BR_index <- which(BR_condition)[1]
BR_index <- FFC_index + BR_index + 3 # will be +1 when at 100Hz

# Index the first frame after the occurrence of BR_index
if (!is.na(BR_index) && BR_index < nrow(data)) {
  data$Events[BR_index] <- 1
   }

rm(BR_condition, max_BR_value)


# End of Follow-Through (EFT) - most posterior position of the hand's COM (X) after BR

# Check if ball release has occurred
if (!is.na(BR_index) && BR_index < nrow(data)) {
  # Subset the data within 100 frames of ball release starting at BR_index
  subset_data <- data[BR_index:(BR_index + 100), ]

   # Find the index of the first peak after ball release
  peak_index <- which(subset_data$`Event - EFT` == max(subset_data$`Event - EFT`, na.rm = TRUE))[1]

  # Check if there is a peak after ball release
  if (!is.na(peak_index)) {
    # Find the index of the first minimum value after the peak
    EFT_index <- which.min(subset_data$`Event - EFT`[peak_index:nrow(subset_data)]) + peak_index - 1
    
    # Mark the row where the first minimum value after the peak occurs with '1' in the original data
    data$Events[BR_index + EFT_index - 1] <- 1
    
  }
}

EFT_index <- BR_index + (EFT_index - 1)

rm(peak_index)

# Check the number of marked events
sum(data$Events)

```


```{r, eval = FALSE Calculate Arm Path}
# Subset data from BR-5 to BR+5 and select just wrist variables

wrist_data <- subset_data %>%
  head(10) %>%
  select(`Wrist Position (Y)`, `Wrist Position (Z)`)

# Linear model for wrist data
wrist_model <- lm(`Wrist Position (Y)` ~ `Wrist Position (Z)`, data = wrist_data)

# Print the summary of the linear model
summary(wrist_model)

# Extract the slope and intercept
slope <- coef(wrist_model)[2]
intercept <- coef(wrist_model)[1]

# Print the slope and intercept
print(paste("Slope:", slope))
print(paste("Intercept:", intercept))

# Scatter plot of the data points and trend line
library(ggplot2)
ggplot(wrist_data, aes(x = `Wrist Position (Z)`, y = `Wrist Position (Y)`)) +
  geom_point(color = "blue", size = 3) +
  labs(title = "Scatter Plot with Trendline",
       x = "Wrist Position (Z)",
       y = "Wrist Position (Y)") +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  theme_minimal()

# Calculate the incline angle in radians
incline_radians <- atan(slope)

# Convert radians to degrees
incline_degrees <- incline_radians * (180 / pi)

# Print the incline angle
print(paste("Incline Angle (degrees):", incline_degrees))

# Convert object into a data frame
arm_path <- data.frame(incline_degrees)
colnames(arm_path) <- "Arm Path"
rownames(arm_path) <- NULL

# Add File Name to arm_path data frame
arm_path <- arm_path %>%
  mutate("File Name" = file_name, .before = "Arm Path")

# Assuming `arm_path` is your data frame and `file_path` contains the path of the imported file
output_folder <- file.path(dirname(file_path), "Arm Path")

# Create the "Arm Path" folder if it doesn't exist
dir.create(output_folder, showWarnings = FALSE)

# Save the data frame to a CSV file in the "Arm Path" folder
write.csv(arm_path, file = file.path(output_folder, paste0(file_name, ".csv")), row.names = FALSE)

rm(wrist_data, arm_path, wrist_model, slope, intercept, incline_radians, incline_degrees, subset_data)

```


```{r Extract Values at Events}
# Create a list of indices and corresponding names
event_indices <- list(
  INI = INI_index,
  FFC = FFC_index,
  MER = MER_index,
  BR = BR_index,
  EFT = EFT_index)

# Create 'At_Events' data frame
At_Events <- data.frame(Event = character(0), stringsAsFactors = FALSE)

# Populate 'At_Events' with index names and corresponding rows
for (event_type in names(event_indices)) {
  index <- event_indices[[event_type]]
  if (!is.na(index)) {
    row_data <- data[index, ]
    row_data$Event <- event_type
    At_Events <- rbind(At_Events, row_data)
  }
}

# Reorder columns to have 'Event' as the first column
At_Events <- At_Events[, c("Event", setdiff(names(At_Events), "Event"))]

# Remove Event and KS columns from final output
At_Events <- At_Events %>%
  select(-Events, -`Event - INI`,  -`Event - FFC`, -`Event - MER`, -`Event - BR`, -`Event - EFT`)
         #-`Pelvis Velocity (KS)`, -`Trunk (KS)`, -`Upper Arm (KS)`, 
         #-`Forearm (KS)`, -`Hand (KS)`)

# Add a column with the file's name
file_name <- tools::file_path_sans_ext(basename(file_path))
At_Events <- cbind(`File Name` = file_name, At_Events)

```

```{r Add Minimum & Maximum Values to 'At Events'}
# Calculate minimum and maximum values for each column in 'data'
min_values <- sapply(data, min, na.rm = TRUE)
max_values <- sapply(data, max, na.rm = TRUE)

# Create 'MinMax' data frame with min and max values
MinMax <- data.frame(rbind(min_values, max_values), check.names = FALSE)
rownames(MinMax) <- NULL # Remove row names

# Add a column with labels "min" and "max"
MinMax$Event <- c("Min", "Max")

# Add a column with the trial name
file_name <- tools::file_path_sans_ext(basename(file_path))
MinMax$`File Name` <- file_name

# Reorder columns to match the original 'data' data frame
MinMax <- MinMax[, c("File Name", "Event", names(data))]

# Remove Event columns
MinMax <- MinMax %>%
  select(-Events,  -`Event - INI`, -`Event - FFC`, -`Event - MER`, -`Event - BR`, -`Event - EFT`)
         #-`Pelvis Velocity (KS)`, -`Trunk (KS)`, -`Upper Arm (KS)`, 
         #-`Forearm (KS)`, -`Hand (KS)`)

# Add minimum and maximum values to bottom of 'At_Events' data frame
At_Events <- rbind(At_Events, MinMax)

rm(max_values, min_values, MinMax)

```

```{r Calculate the % of min and max indexes}
# Calculate the range between INI_index and BR_index
range <- BR_index - INI_index + 1

# Calculate the percentage each frame represents
frame_perc <- 100 / range

# Initialize dataframes to store minimum and maximum values
min_df <- data.frame(matrix(nrow = 1, ncol = 15))
max_df <- data.frame(matrix(nrow = 1, ncol = 15))

# Specify the range of columns
columns_of_interest <- 1:15

# Loop through each column
for (col_index in columns_of_interest) {
  col_name <- names(data)[col_index]  # Get the original column name
  
  # Find index of minimum and maximum values between INI_index and EFT_index
  min_index <- which.min(data[INI_index:EFT_index, col_index]) + INI_index - 1
  max_index <- which.max(data[INI_index:EFT_index, col_index]) + INI_index - 1
  
  # Calculate the percentage indices
  min_percentage_index <- (min_index - INI_index) * frame_perc
  max_percentage_index <- (max_index - INI_index) * frame_perc
  
  # Store the results in the dataframes
  min_df[1, col_index] <- min_percentage_index
  max_df[1, col_index] <- max_percentage_index
}

# Set column names
colnames(min_df) <- colnames(max_df) <- names(data)

percentage <- rbind(min_df, max_df)
rownames(percentage) <- NULL # Remove row names

# Add a column with labels "min" and "max"
percentage$Event <- c("Min % Index", "Max % Index")

# Add a column with the trial name
file_name <- tools::file_path_sans_ext(basename(file_path))
percentage$`File Name` <- file_name

# Reorder columns to match the original 'data' data frame
percentage <- percentage[, c("File Name", "Event", names(data))]

# Remove Event columns
percentage <- percentage %>%
  select(-Events,  -`Event - INI`, -`Event - FFC`, -`Event - MER`, -`Event - BR`, -`Event - EFT`)

At_Events <- rbind(At_Events, percentage)

rm(min_df, max_df, percentage, subset_data, col_index, col_name, columns_of_interest, frame_perc,
   max_index, min_index, max_percentage_index, min_percentage_index, range)

```


```{r}
# Save 'At_Events' as a CSV file in the same directory as the original .txt file
write.csv(At_Events, file = file.path(dirname(file_path), paste0(file_name, " (at events).csv")), row.names = FALSE)

# Print a confirmation message
print(At_Events[, 1:3])

# Clean Console
rm(INI_index, FFC_index, MER_index, BR_index, EFT_index, event_type, event_indices, index, row_data)

```

```{r}
# Save 'data' as a tab-delimited text file, overwriting the original .txt file
write.table(data, file = file_path, sep = "\t", row.names = FALSE, col.names = TRUE)

# Print a confirmation message
print(paste("Data saved to:", file_path))

```

# Once I have event marked trials, I will look to determine the duration between events (phases).

# I also want to find the angular displacement across phases (current value - previous value).
