# This code event marks pitching trials, then trims the file from initiation (INI) 
# to end of follow-through (EFT). Once the file has been event marked, it extracts
# shoulder and elbow kinetics (shoulder IR torque, shoulder horizontal adduction
# torque, and elbow varus torque), and temporal sequence of 5 segments (pelvis,
# shoulder, elbow, wrist, and finger). Sequence values and indices are saved too.

```{r Load the required packages, eval = FALSE}
library(tidyverse)

```

```{r Import the raw trial file (.txt), extract the file name, and remove the last empty column}
# Let the user select the file to import
file_path <- file.choose()

# Extract the file name from the file path
file_name <- tools::file_path_sans_ext(basename(file_path))
# file_end <- substr(file_name, nchar(file_name) - 9, nchar(file_name))
# file_name <- sub("(.*mph).*", "\\1", file_name)

# Read the data
df <- read.delim(file_path, header = TRUE, skip = 8, check.names = FALSE)

# # Remove the final column in the data if it's empty
df <- df %>%
  select(- 28) 

```

```{r Define the sampling rate, then initialise a column for the events, and time}
# Define sampling rate (Hz)
SampleRate <- 238.095 # change as necessary

# Create and events column and populate with 0s
df$Events <- 0

# Create the column "Time (s)" (divide by sampling rate)
df$`Time (s)` <- df$`Sample #` / SampleRate

df <- df %>%
  select(Events, `Time (s)`, everything())

```

```{r Define Event - FFC}
Event_FFC <- which(diff(df$`Event - FFC`) > 20)[1] # adjust threshold as necessary
plot(df$`Event - FFC`)

```

```{r Define Event - INI}
Event_INI <- which(df$`Event - INI`[(Event_FFC-360):nrow(df)] > 0.2)[1] # adjust threshold as necessary
Event_INI <- (Event_FFC - 360) + Event_INI
plot(df$`Event - INI`)

```

```{r Define Event - MAW}
Event_MAW <- which(df$`Event - MAW` == min(df$`Event - MAW`[(Event_FFC - 100):nrow(df)], na.rm = TRUE))[1]
plot(df$`Event - MAW`)

```


```{r Define Event - BR}
Event_BR <- which(df$`Event - BR` == max(df$`Event - BR`, na.rm = TRUE))[1]
plot(df$`Event - BR`)

```
```{r Define Event - MER}
Event_MER <- which(df$`Event - MER` == max(df$`Event - MER`[Event_MAW:Event_BR], na.rm = TRUE))[1]
plot(df$`Event - MER`)

```

```{r Define Event - EFT}
# Check if ball release has occurred
if (!is.na(Event_BR) && Event_BR < nrow(df)) {
  # Subset the data from ball release to ball release +100
  subset_data <- df[(Event_BR + 10):(Event_BR + 100), ]

  # Find the index of the row where FT is at its smallest in the subset
  EFT_index <- which(subset_data$`Event - EFT` == min(subset_data$`Event - EFT`, na.rm = TRUE))[1]
  
}

Event_EFT <- (Event_BR + 10) + EFT_index # determine event index in full data frame
plot(df$`Event - EFT`)

```

```{r Mark each event index in "data" with a '1'}
# Mark Event_BFC with '1'
if (!is.na(Event_INI)) {
  df$Events[Event_INI] <- 1
}

# Mark Event_FFC with '1'
if (!is.na(Event_FFC)) {
  df$Events[Event_FFC] <- 1
}

# Mark Event_MAW with '1'
if (!is.na(Event_MAW) && Event_MAW < nrow(df)) {
  df$Events[Event_MAW] <- 1
}

# Mark Event_MER with '1'
if (!is.na(Event_MER) && Event_MER < nrow(df)) {
  df$Events[Event_MER] <- 1
}

# Mark Event_BR + 2 with '1'
if (!is.na(Event_BR) && Event_BR < nrow(df)) {
  df$Events[Event_BR + 2] <- 1
}

# Mark Event_EFT with '1'
if (!is.na(Event_EFT)) {
    df$Events[Event_EFT] <- 1
}

# Check the number of marked events
sum(df$Events)

```

```{r Trim the file from INI to EFT, then save}
trimmed <- df[Event_INI:Event_EFT, ] %>%
  select(25:29)

# Specify the folder to save the data to
output_folder <- "C:/Users/billy/OneDrive/Documents/Academics/2. Auburn/Sports Medicine & Movement Lab/Misc/Caroline Keller/Fellowship/Data/Trimmed Trials (INI to EFT)/Temporal Data"

# Create the folder if it doesn't exist
dir.create(output_folder, showWarnings = FALSE)

write.csv(trimmed, file = file.path(output_folder, paste0(file_name, ".csv")), row.names = FALSE)

```

```{r Save event indices as a .csv file, with the extracted file name}
event_indices <- cbind(file_name, Event_INI, Event_FFC, Event_MAW, Event_MER, Event_BR, Event_EFT)
colnames(event_indices) <- c("File Name", "Event - INI", "Event - FFC", "Event - MAW",
                             "Event - MER", "Event - BR", "Event - EFT")
rownames(event_indices) <- NULL

# Specify the folder to save the data to
output_folder <- "C:/Users/billy/OneDrive/Documents/Academics/2. Auburn/Sports Medicine & Movement Lab/Misc/Caroline Keller/Fellowship/Data/Event Indices/Temporal Data"

# Create the folder if it doesn't exist
dir.create(output_folder, showWarnings = FALSE)

write.csv(event_indices, file = file.path(output_folder, paste0(file_name, ".csv")), row.names = FALSE)

print(event_indices)

```

```{r Extract arm kinetics between MAW and BR, eval = FALSE}
df_kinetics <- df[Event_MAW:Event_BR, ]

shoulder_IR_torque <- min(df_kinetics$`Shoulder Rotation Torque (Nm)`)
shoulder_AP_torque <- min(df_kinetics$`Shoulder AP Torque (Nm)`)
elbow_ML_torque <- min(df_kinetics$`Elbow ML Torque (Nm)`)
# wrist_AP_torque <- min(df_kinetics$`Wrist AP Torque (Nm)`)

```

```{r Extract arm kinematics, eval = FALSE}
# Joint Angles
shoulder_Rot <- max(trimmed$`Shoulder Rotation`)
elbow_flexion <- max(trimmed$`Elbow Fle/Ext`)
wrist_extension <- min(trimmed$`Wrist Fle/Ext`)
wrist_pronation_BR <- trimmed$`Wrist Pro/Sup`[trimmed$Sample == Event_BR - 3]

# Joint Velocities
shoulder_Rot_vel <- min(trimmed$`Shoulder Rotation Velocity`)
elbow_extension_vel <- min(trimmed$`Elbow Fle/Ext Velocity`)
wrist_flexion_vel <- max(trimmed$`Wrist Fle/Ext Velocity`)
wrist_pronation_vel <- trimmed$`Wrist Pro/Sup Velocity`[trimmed$Sample == Event_BR - 3]

```

```{r Extract peak linear velocities and indices for sequencing variables}
# Linear Pelvis Velocity
pelvis_x_vel <- max(df$`Linear Velocity - Pelvis`[Event_INI:Event_BR])
pelvis_idx <- which.max(df$`Linear Velocity - Pelvis`[Event_INI:Event_BR])

# Linear Shoulder Velocity
shoulder_x_vel <- max(df$`Linear Velocity - Shoulder`[Event_INI:Event_BR])
shoulder_idx <- which.max(df$`Linear Velocity - Shoulder`[Event_INI:Event_BR])

# Linear Elbow Velocity
elbow_x_vel <- max(df$`Linear Velocity - Elbow`[Event_INI:Event_BR])
elbow_idx <- which.max(df$`Linear Velocity - Elbow`[Event_INI:Event_BR])

# Linear Wrist Velocity
wrist_x_vel <- max(df$`Linear Velocity - Wrist`[Event_INI:Event_BR])
wrist_idx <- which.max(df$`Linear Velocity - Wrist`[Event_INI:Event_BR])

# Linear Finger Velocity
finger_x_vel <- max(trimmed$`Linear Velocity - Finger`)
finger_idx <- which.max(trimmed$`Linear Velocity - Finger`)

sequence <- cbind(file_name, pelvis_x_vel, pelvis_idx, shoulder_x_vel, shoulder_idx,
                  elbow_x_vel, elbow_idx, wrist_x_vel, wrist_idx,
                  finger_x_vel, finger_idx)

colnames(sequence) <- c("File Name", "Linear Velocity - Pelvis", "Linear Velocity - Pelvis (idx)",
                        "Linear Velocity - Shoulder", "Linear Velocity - Shoulder (idx)",
                        "Linear Velocity - Elbow", "Linear Velocity - Elbow (idx)",
                        "Linear Velocity - Wrist", "Linear Velocity - Wrist (idx)",
                        "Linear Velocity - Finger", "Linear Velocity - Finger (idx)")
```

```{r Visualise the sequence of peak linear velocities}
plot_data <- data.frame(variable = factor(c("pelvis_idx", "shoulder_idx", "elbow_idx", "wrist_idx", "finger_idx"),
                                          levels = c("pelvis_idx", "shoulder_idx", "elbow_idx", "wrist_idx", "finger_idx")),
                        value = as.numeric(sequence[, c("Linear Velocity - Pelvis (idx)", "Linear Velocity - Shoulder (idx)", "Linear Velocity - Elbow (idx)", "Linear Velocity - Wrist (idx)", "Linear Velocity - Finger (idx)")]))

ggplot(plot_data, aes(x=value, y=variable)) +
  geom_point()

```


```{r plot the velocity waveforms}
ggplot(df[Event_INI:Event_EFT, ], aes(x = `Sample #`)) +
  geom_line(aes(y = `Linear Velocity - Pelvis`, color = "Linear Velocity - Pelvis")) +
  geom_line(aes(y = `Linear Velocity - Shoulder`, color = "Linear Velocity - Shoulder")) +
  geom_line(aes(y = `Linear Velocity - Elbow`, color = "Linear Velocity - Elbow")) +
  geom_line(aes(y = `Linear Velocity - Wrist`, color = "Linear Velocity - Wrist")) +
  geom_line(aes(y = `Linear Velocity - Finger`, color = "Linear Velocity - Finger"))

```

```{r Combine kinetics, kinematics, and temporal sequence, then save, eval = FALSE}
df_combined <- cbind(file_name, shoulder_IR_torque, shoulder_AP_torque, elbow_ML_torque,
                     wrist_AP_torque, shoulder_Rot, shoulder_Rot_vel, elbow_flexion,
                     elbow_extension_vel, wrist_extension, wrist_flexion_vel, wrist_pronation_BR,
                     wrist_pronation_vel)

colnames(df_combined) <- c("File Name", "Shoulder Rotation Torque (Nm)", "Shoulder Horizontal Adduction Torque (Nm)", "Elbow Varus Torque (Nm)", "Wrist Flexion Torque (Nm)", "Max Shoulder ER", "Max IR Velocity", "Max Elbow Flexion", "Max Elbow Extension Velocity", "Max Wrist Extension", "Max Wrist Flexion Velocity", "Wrist Rotation at BR", "Wrist Rotation Velocity at BR")

# df_combined <- cbind(df_combined, sequence)

# Specify the folder to save the data to
output_folder <- "C:/Users/billy/OneDrive/Documents/Academics/2. Auburn/Sports Medicine & Movement Lab/Misc/Caroline Keller/Fellowship/Data/Kinetics & Kinematics"

# # Specify the folder to save the data to (use when sequence variables are available)
# output_folder <- "C:/Users/billy/OneDrive/Documents/Academics/2. Auburn/Sports Medicine & Movement Lab/Misc/Caroline Keller/Fellowship/Data/Kinetics, Kinematics, and Temporal Sequence"

# Create the folder if it doesn't exist
dir.create(output_folder, showWarnings = FALSE)

write.csv(df_combined, file = file.path(output_folder, paste0(file_name, ".csv")), row.names = FALSE)

```

```{r Save sequence as .csv file}
# Specify the folder to save the data to
output_folder <- "C:/Users/billy/OneDrive/Documents/Academics/2. Auburn/Sports Medicine & Movement Lab/Misc/Caroline Keller/Fellowship/Data/Kinetics & Kinematics/Sequence Data"

# Create the folder if it doesn't exist
dir.create(output_folder, showWarnings = FALSE)

write.csv(sequence, file = file.path(output_folder, paste0(file_name, ".csv")), row.names = FALSE)

```

